---
title: "Stats 2 Project 2"
author: "Rick Farrow, Bruce Granger, TQ Senkungu"
date: "12/1/2018"
output:
    html_document:
        keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Required Libraries
- TIDYVERSE
```{r Load_Libraries, include=FALSE}
library(tidyverse)
```

## Load Data

- We have some Breast Cancer Data from the Wisconsin Diagnostic Breast Cancer (WDBC) dataset and will first load it in.
- Add an attribute called CancerState, which is similiar to the Class attribute, the difference is that CancerState uses words to describe the condition, meaning when Class == 2, then "Benign" and when Class == 4, then "Malignanat"

```{r input_data}
wdbc_data <- read.csv("https://raw.githubusercontent.com/tikisen/6372_proj2/master/Data/breast-cancer-wisconsin-data.csv", 
                      sep = ",", 
                      row.names = NULL, 
                      header = TRUE)

entire.dataset <- wdbc_data %>% mutate(CancerState = case_when(Class == 2 ~ "Benign",
                                                                    Class == 4 ~ "Malignanat"))
entire.dataset$CancerState <- as.factor(entire.dataset$CancerState)
```

# Bruce's Work Starts Here

## Summary Statistics/Histograms
```{r SummaryStats}
entire.dataset %>% select(c(2:12)) %>% summary() 

# TOTAL COUNT BY CANCER TYPE ####
ggplot(data=entire.dataset, aes(x=CancerState, colour = CancerState)) +
  geom_bar() +
  geom_text(stat='Count', aes(label=..count..), vjust = 10) +
  theme(legend.position = "none") +
  scale_color_manual(values = c("blue", "red")) +
  ggtitle("Total Count by Cancer Type: Blue = Benign; Red = Malignanat") 
  

# PERCENT OF TOTAL BY CANCER TYPE ####
entire.dataset.percent <- entire.dataset %>% 
  count(CancerState) %>% 
  mutate(perc = n / nrow(entire.dataset))

ggplot(data=entire.dataset.percent, aes(x = CancerState, y = perc, colour = CancerState)) +
  geom_bar(stat = "identity") +
  geom_text(stat = "identity", aes(label=round(perc*100,2)), vjust = 10) +
  theme(legend.position = "none") +
  scale_color_manual(values = c("blue", "red")) +
  ggtitle("Percent of Total by Cancer Type: Blue = Benign; Red = Malignanat")
  
```

## Pairs Plots
- Version 1:
- The objective of the "Pairs Plot" is to create plots based upon paring of variables
- Additionaly, each observation is color coded to simutaleniously see if the observation is "Benign" or "Malignanat" cancer.  
  - Color Coding:
    * Blue = Benign 
    * Red = Malignanat

```{r pairsplot_ver1}
ed.small <- entire.dataset %>% select(-c(1,12))
cols <- character(nrow(ed.small))
cols[] <- "black"
cols[ed.small$Class == 2] <- "blue"
cols[ed.small$Class == 4] <- "red"
pairs(ed.small, col=cols, main = "WDBC Pairs Plot: Blue = Benign; Red = Malignanat")
```
- Version 2:
- Similar to the first Pairs Plot, however version 2 introduces jitter to the observations and as a result it is easy to see the density of the observations.
```{r pairsplot_ver2}
# CONVERT TO NUMERIC
ed.small$Clump_Thickness <- as.integer(ed.small$Clump_Thickness)
ed.small$Uniformity_Cell_Size <- as.integer(ed.small$Uniformity_Cell_Size)
ed.small$Uniformity_Cell_Shape <- as.integer(ed.small$Uniformity_Cell_Shape)
ed.small$Marginal_Adhesion <- as.integer(ed.small$Marginal_Adhesion)
ed.small$Single_Epithelial_Cell_Size <- as.integer(ed.small$Single_Epithelial_Cell_Size)
ed.small$Bare_Nuclei <- as.integer(ed.small$Bare_Nuclei)
ed.small$Bland_Chromatin <- as.integer(ed.small$Bland_Chromatin)
ed.small$Normal_Nucleoli <- as.integer(ed.small$Normal_Nucleoli)
ed.small$Mitoses <- as.integer(ed.small$Mitoses)
ed.small$Class <- as.integer(ed.small$Class)

# REDUCE LENGTH OF ATTRIBUTES 
ed.small <- ed.small %>% select(Thick = Clump_Thickness, U_Size = Uniformity_Cell_Size,
                                Shape = Uniformity_Cell_Shape, Adhesion = Marginal_Adhesion,
                                Epi_Size = Single_Epithelial_Cell_Size, Bare = Bare_Nuclei,
                                Bland = Bland_Chromatin, Nucleoli = Normal_Nucleoli, 
                                Mitoses = Mitoses, Class = Class)

# the alpha argument in rgb() lets you set the transparency
cols2 = c(rgb(red=0, green=0, blue=255, alpha=50, maxColorValue=255), 
          rgb(red=255, green=0, blue=0, alpha=50, maxColorValue=255))
cols2 = ifelse(ed.small$Class==2, cols2[1], cols2[2])

# here we jitter the data
set.seed(6141)  # this makes the example exactly reproducible
jbreast = apply(ed.small[,1:9], 2, FUN=function(x){ jitter(x, amount=.5) })
jbreast = cbind(jbreast, class=ed.small[,10])  # the class variable is not jittered

#windows()  # to match up the 1st & 2nd sets requires more coding
layout(matrix(1:25, nrow=5, byrow=T))
par(mar=c(.5,.5,.5,.5), oma=c(2,2,2,2))

for(i in 1:5){
   for(j in 6:10){
    
    plot(jbreast[,j], jbreast[,i], col=cols2, pch=16,
         
         axes=F, main="", xlab="", ylab="")
    
    box()
    
    if(j==6 ){ mtext(colnames(jbreast)[i], side=2, cex=.7, line=1) }
    
    if(i==5 ){ mtext(colnames(jbreast)[j], side=1, cex=.7, line=1) }
    
    if(j==10){ axis(side=4, seq(2,10,2), cex.axis=.8) }
    
    if(i==1 ){ axis(side=3, seq(2,10,2), cex.axis=.8) }
    
  }
  
}

```
## Histogram
- The histograms are a look at each attribute broken down by CancerState
```{r histogram}
ed.small <- entire.dataset %>% select(-c(1,11))

ed.small$Clump_Thickness <- as.integer(ed.small$Clump_Thickness)
ed.small$Uniformity_Cell_Size <- as.integer(ed.small$Uniformity_Cell_Size)
ed.small$Uniformity_Cell_Shape <- as.integer(ed.small$Uniformity_Cell_Shape)
ed.small$Marginal_Adhesion <- as.integer(ed.small$Marginal_Adhesion)
ed.small$Single_Epithelial_Cell_Size <- as.integer(ed.small$Single_Epithelial_Cell_Size)
ed.small$Bare_Nuclei <- as.integer(ed.small$Bare_Nuclei)
ed.small$Bland_Chromatin <- as.integer(ed.small$Bland_Chromatin)
ed.small$Normal_Nucleoli <- as.integer(ed.small$Normal_Nucleoli)
ed.small$Mitoses <- as.integer(ed.small$Mitoses)

ed.tall <- ed.small %>% 
    select(Clump = Clump_Thickness, Cell_Size = Uniformity_Cell_Size, 
         Cell_Shape = Uniformity_Cell_Shape, Adhesion = Marginal_Adhesion,
         Epithelial = Single_Epithelial_Cell_Size, Nuclei = Bare_Nuclei,
         Chromatin = Bland_Chromatin, Nucleoli = Normal_Nucleoli, everything()) %>%
  gather(-10, key = "Variable", value = "Value") %>% 
  filter(!is.na(CancerState))

ggplot(data = ed.tall, aes(x=Value)) +
  geom_histogram(bins=15) + 
  facet_wrap(Variable ~ CancerState, ncol = 9)
```

# Bruce's Work Ends Here

# Rick's Work Starts Here

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
library(MASS)
ftable(addmargins(table(wdbc_data$Class,wdbc_data$Mitoses)))
prop.table(table(wdbc_data$Class,wdbc_data$Mitoses),2)
```

# Rick's Work Ends Here

#TQ's Work Starts Here

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


#TQ's Work Ends Here